Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-11 write-ups/PLDI/paper.tex
Add better discussion with line references in cooperative version of work stealing

\begin{lstlisting}
(*@\label{line:wksteal:kernelfunc}@*) kernel work_stealing(global Task * pools) {
  while (more_work(pools)) {
(*@\label{line:wksteal:getgroupid}@*)    int pool_id = get_group_id();
(*@\label{line:wksteal:poporsteal}@*)    Task * t = pop_or_steal(pools, pool_id);
    if (t) {
      // task processing may create more work
(*@\label{line:wksteal:processtask}@*)      process_task(t, pools, pool_id);
    }
  }
}
\end{lstlisting}

@@

\myfiglong\ref{fig:workstealing-cooperative} shows a cooperative
version of the work stealing example of \myfig\ref{fig:workstealing} that we
discussed in \mysec\ref{sec:openclexamples}.
\ADComment{TODO: go through with Hugues; where are the offer functions?}

For the work stealing example, there is no state to transmit since a
computation is entirely parameterised by a task, which is retrieved from
a task pool located in global memory. Care must be taken on the dynamic
change of the number of workgroups though: we cannot assume that the
total number of task pools is equal to $\getnumgroups$. Thus, the number
of pools must be given as an argument, and the $\mathsf{pool}$ argument
must be obtained from the group id modulo the total number of
pools. This total number of pools must also be passed to
$\mathsf{pop\_or\_steal}$, since when a workgroup wants to steal it must
access other pool indexes, which are bounded by the total number of
pools.

@@

\begin{lstlisting}
kernel work_stealing(global Task *pools, const int num_pools) {
  while (more_work(pools)) {
    int pool_id = get_group_id() % num_pools;
    Task *t = pop_or_steal(pools, pool_id, num_pools);
    if (t) {
      // task processing may create more work
      process_task(t, pools, pool_id);
    }
  }
}
\end{lstlisting}

>>>>

\begin{lstlisting}
(*@\label{line:wksteal:kernelfunc}@*)kernel work_stealing(global Task * pools) {
(*@\label{line:wksteal:getgroupid}@*)  int pool_id = get_group_id();
  while (more_work(pools)) {
(*@\label{line:wksteal:poporsteal}@*)    Task * t = pop_or_steal(pools, pool_id);
    if (t) {
      // task processing may create more work
(*@\label{line:wksteal:processtask}@*)      process_task(t, pools, pool_id);
    }
  }
}
\end{lstlisting}

@@

\myfiglong\ref{fig:workstealing-cooperative} shows a cooperative
version of the work stealing example of \myfig\ref{fig:workstealing} that we
discussed in \mysec\ref{sec:openclexamples}.

For the work stealing example, there is no state to transmit since a
computation is entirely parameterised by a task, which is retrieved from
a task pool located in global memory. In the main loop, before trying to
obtain a task, we make the workgroup offer itself to be forked or killed
(lines~\ref{line:cwksteal:offerfork},\ref{line:cwksteal:offerkill}). Note
that a workgroup may be killed even if its associated task pool is not
empty, since remaining tasks could eventually be stealed by other
workgroups.  Care must be taken on computing pool identifiers: the
original version in \myfig\ref{fig:workstealing} can rely on the fact
that there are as many pools as workgroups, but here the number of
workgroups may change during execution. Therefore, the number of pools
must now be explicitely given as an argument
(line~\ref{line:cwksteal:argnumpool}) and used as an upper bound of pool
indexes when scanning pools in $\mathsf{more\_work}$
(line~\ref{line:cwksteal:morework}) and $\mathsf{pop\_or\_steal}$
(line~\ref{line:cwksteal:poporsteal}). Besides, computation of the
workgroup's pool index (line~\ref{line:cwksteal:getgroupid}) must not
only be bounded by the number of pools but also be moved inside the main
loop since a forked workgroup that starts to execute after $\offerfork$
must re-compute its pool index.

@@

\begin{lstlisting}
kernel work_stealing(global Task * pools,
(*@\label{line:cwksteal:argnumpool}@*)                     const int num_pools) {
(*@\label{line:cwksteal:morework}@*)  while (more_work(pools, num_pools)) {
(*@\label{line:cwksteal:offerfork}@*)    offer_fork();
(*@\label{line:cwksteal:offerkill}@*)    offer_kill();
(*@\label{line:cwksteal:getgroupid}@*)    int pool_id = get_group_id() % num_pools;
(*@\label{line:cwksteal:poporsteal}@*)    Task * t = pop_or_steal(pools, pool_id, num_pools);
    if (t) {
      // task processing may create more work
(*@\label{line:cwksteal:processtask}@*)      process_task(t, pools, pool_id);
    }
  }
}
\end{lstlisting}
